Analysis paper

### 2021-03-31 analysis/analysis.Rmd
update analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(covid.german.forecasts)
library(here)
library(magrittr)
library(scoringutils)
library(knitr)
library(kableExtra)
library(data.table)
library(ggplot2)

@@

## Figure with 2 week ahead forecasts and scores

- [x] create plots
- [x] combine plots into one
- [ ] shade dates that are excluded in scores
- [ ] maybe add density plot on the y axis
- [ ] maybe use mean scores ratio instead of WIS for plot with score over time

### Forecasts vs. truth data
```{r}
df <- combined_data[as.Date(forecast_date) %in% forecast_dates$unfiltered &
                      grepl("2", target) & 
                      quantile %in% c(0.025, 0.25, 0.75, 0.975)]

@@

df %>%
  ggplot(aes(y = true_value, x = target_end_date)) + 
  geom_linerange(aes(ymin = `0.025`, ymax = `0.975`, color = model), 
                 size = 1.5,
                 alpha = 0.4,
                 position = position_dodge(width = 3)) + 
  geom_linerange(aes(ymin = `0.25`, ymax = `0.75`, color = model), 
                 size = 1.5,
                 alpha = 0.8,
                 position = position_dodge(width = 3)) + 
  labs(y = "value", x = "date") + 
  geom_point() + 
  geom_line() + 
  theme_light() + 
  facet_grid(target_type ~ location_name, scales = "free") + 
  theme(legend.position = "bottom")

ggsave(here("analysis", "plots", "truth-and-2-wk-ahead.png"))

@@

@@

- severe overshooting for Renewal model
- need to increase value for position dodge
- reporting artifacts around christmas

### plot with interval score

```{r}
# plot with scores over time ---------------------------------------------------
scores_target_forecastdate <- eval_forecasts(
  data = data[grepl("2", target)],
  summarise_by = c("model", "location_name", "target_type", "forecast_date"), 
  compute_relative_skill = TRUE, 
  baseline = "Baseline"
)

scores_target_forecastdate  %>%
  ggplot(aes(x = as.Date(forecast_date), y = interval_score, colour = model)) + 
  geom_line() + 
  geom_point() + 
  theme_light() + 
  facet_grid(target_type ~ location_name, scales = "free") +
  theme(legend.position = "bottom") +
  labs(y = "weighted interval score", x = "date")

ggsave(here("analysis", "plots", "wis-over-time.png"))
```

- all models perform terribly around Christmas
- the effect is much stronger in Germany than in Poland
- these artifacts around Christmas hit Renewal hardest
- in Germany scores are high in the beginning and the end, but it doesn't seem to correlate perfectly with the absolute value of the number of weekly cases

@@

- maybe look at correlation between the number of weekly cases and the score?
- look at scores for rising and falling parts of the epidemic separately

### same plot with mean score ratio
```{r}
scores_target_forecastdate  %>%
  ggplot(aes(x = as.Date(forecast_date), y = scaled_rel_skill, colour = model)) + 
  geom_line() + 
  geom_point() + 
  theme_light() + 
  facet_grid(target_type ~ location_name, scales = "free") +
  theme(legend.position = "bottom") +
  labs(y = "mean score ratio", x = "date") 
```

- There is a gap in the plot where there shouldn't be a gap. Nikos is a bad Padawan and needs to look at this again. Need to think a minute about what we actually see here. It is not exactly the WIS of one model over the WIS of the baseline, but something slightly different
- In German cases there are two spikes that I think would be interesting to look at
- death forecasts look quite a lot messier than case forecasts

@@

- [x] create summary table for scores visualised before
- [x] exclude Christmas period
- [x] use whole period for cases
- [x] use only period for which we have all models for deaths
- [] also add 50% and 90% coverage in table

@@

@@

@@

- one possibility would be to take the ridge plot from Cramer et al from the Hub
- decide on whether to use rank or standardised rank

@@

library(ggridges)
ggplot(ranked_scores, aes(y = model, x = standard_rank, 
                          group = model, fill = model)) + 
  ggridges::geom_density_ridges()

ggplot(ranked_scores, aes(y = model, x = standard_rank, 
                          group = model, fill = model)) + 
  ggridges::geom_density_ridges() + 
  facet_wrap(~ target_type)

ggplot(ranked_scores, aes(y = model, x = standard_rank, 
                          group = model, fill = model)) + 
  ggridges::geom_density_ridges() + 
  facet_wrap(location_name ~ target_type)

@@

- bimodal distribution is interesting
- maybe a ridge plot is not a good way of analysing this as values go from -50 to 150 on the plot? In reality, they should be between 0 and 100. Is there an option to adapt this?

@@

wis_components(scores_target, 
               facet_formula = ~ target_type, 
               scales = "free_x", 
               x_text_angle = 0) + 
  theme(legend.position = "bottom") + 
  coord_flip() + 
  labs(title = "Components of the weighted interval score for 2-week ahead forecasts", 
       x = NULL)

@@

scores_target <- eval_forecasts(
  data[grepl("2", target)], 
  summarise_by = c("model", "location_name", "target_type"), 
  compute_relative_skill = TRUE
)

@@

```{r, eval = FALSE}
scores_target <- eval_forecasts(
  data[grepl("2", target)], 
  summarise_by = c("model", "target_type", "range", "quantile"), 
  compute_relative_skill = TRUE
)

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(covid.german.forecasts)
library(here)
library(magrittr)
library(scoringutils)
library(knitr)
library(kableExtra)
library(data.table)
library(ggplot2)
library(ggridges)
library(patchwork)

@@

## Visualisation of forecasts and scores

This includes all available dates for the forecast plot. For the score plot it includes the full set for cases, and a restricted set for deaths. The restricted set is all forecasts from "2020-12-07" on to make sure that we have a full set of forecasts for deaths. The plot also includes the Christmas period - it probably makes sense to gray that out.

```{r fig.height=10, fig.width=15}
df <- combined_data[as.Date(forecast_date) %in% forecast_dates$unfiltered &
                      grepl("2", target) & 
                      quantile %in% c(0.025, 0.25, 0.75, 0.975)]

@@

truth_vs_forecast_2 <- df %>%
  ggplot(aes(y = true_value, x = target_end_date)) + 
  geom_linerange(aes(ymin = `0.025`, ymax = `0.975`, color = model), 
                 size = 1.5,
                 alpha = 0.4,
                 position = position_dodge(width = 3)) + 
  geom_linerange(aes(ymin = `0.25`, ymax = `0.75`, color = model), 
                 size = 1.5,
                 alpha = 0.8,
                 position = position_dodge(width = 3)) + 
  labs(y = "value", x = "date") + 
  geom_point() + 
  geom_line() + 
  theme_light() + 
  facet_grid(target_type ~ location_name, scales = "free") + 
  theme(legend.position = "bottom")

ggsave(here("analysis", "plots", "truth-and-2-wk-ahead.png"), 
       plot = truth_vs_forecast_2)

@@

```{r fig.height=10, fig.width=15}
# plot with scores over time ---------------------------------------------------
scores_target_forecastdate <- eval_forecasts(
  data = data[grepl("2", target)],
  summarise_by = c("model", "location_name", "target_type", "forecast_date"), 
  compute_relative_skill = TRUE, 
  baseline = "Baseline"
)

wis_scores_time_2 <- scores_target_forecastdate  %>%
  ggplot(aes(x = as.Date(forecast_date), y = interval_score, colour = model)) + 
  geom_line() + 
  geom_point() + 
  theme_light() + 
  facet_grid(target_type ~ location_name, scales = "free") +
  theme(legend.position = "bottom") +
  labs(y = "weighted interval score", x = "date")

ggsave(here("analysis", "plots", "wis-over-time.png"), 
       plot = wis_scores_time_2)
```

```{r}
### same plot with mean score ratio
rel_skill_time_2 <- scores_target_forecastdate  %>%
  ggplot(aes(x = as.Date(forecast_date), y = scaled_rel_skill, colour = model)) + 
  geom_line() + 
  geom_point() + 
  theme_light() + 
  facet_grid(target_type ~ location_name, scales = "free") +
  theme(legend.position = "bottom") +
  labs(y = "relative skill", x = "date")

ggsave(here("analysis", "plots", "wis-over-time.png"), 
       plot = rel_skill_time_2)
```

```{r fig.height=10, fig.width=20}
# glue two plots together
truth_vs_forecast_2 + rel_skill_time_2
ggsave(here("analysis", "plots", "figure-forecasts-2.png"))
```

@@

- severe overshooting for Renewal model
- all models are terrible around Christmas
- the artifacts around Christmas hit Renewal hardest
- the Christmas effect is much stronger in Germany than in Poland
- need to increase value for position dodge
- reporting artifacts around christmas

**Todo**

- [ ] shade dates that are excluded in "official scoring"
- [ ] maybe add density plot with scores on the y axis
- [ ] remove baseline from relative skill plot
- [ ] make both plots nicer
    - [ ] increase dodge
    - [ ] change colour scheme
    - [ ] adjust size of points
- [ ] presumably change how plots are arranged
- [ ] make final decision whether we want to have relative skill or just WIS on the other plot
    - WIS would show the overall level (which scales with the number of cases)
    - relative skill corrects for that in some sene, but also introduces other biases

@@

- maybe look at correlation between the number of weekly cases and the score?
- look at scores for rising and falling parts of the epidemic separately
- maybe it would be interesting to look at other baseline models, e.g. "how hard is it to beat a model that just always continues the trend?"

@@

This excludes all forecasts made on c("2020-12-21", "2020-12-28").

@@

**To do**
- [] also add 50% and 90% coverage in table
- [] make a final decision (and document it somewhere) about how we want to exclude the Christmas period. Right now we are excluding forecast dates, but do not restrict target end dates in any way.

**Further things to look at**
- maybe think about some kind of stratification by time?

@@

--> not sure what we wanted to do here.

@@

@@

setDT(ranked_scores)
df <- ranked_scores[, .(rank_count = .N), 
                    by = c("target_type", "location_name", "model", "rank")][, 
                      sum := .N, by = c("target_type", "location_name")
                    ][, rank_count := rank_count / sum]

ggplot(df, aes(y = model, x = rank, 
                          group = model, fill = model)) + 
  ggridges::geom_ridgeline(aes(height = rank_count)) + 
  facet_wrap(location_name ~ target_type, scales = "free_x")

ggplot(ranked_scores, aes(y = model, x = rank, 
                          group = model, fill = model)) + 
  ggridges::geom_density_ridges(stat = "binline") + 
  facet_wrap(location_name ~ target_type, scales = "free_x")

@@

- bimodal distribution is interesting

**To do**

- [ ] make plot nicer

@@

wis_components(scores_target, 
               facet_formula = ~ target_type, 
               scales = "free_x", 
               relative_contributions = TRUE,
               x_text_angle = 0) + 
  theme(legend.position = "bottom") + 
  coord_flip() + 
  labs(title = "Components of the weighted interval score for 2-week ahead forecasts", 
       x = NULL)

@@

scores_target <- eval_forecasts(
  data[grepl("4", target)], 
  summarise_by = c("model", "location_name", "target_type"), 
  compute_relative_skill = TRUE
)

@@

```{r, eval = TRUE}
scores_target <- eval_forecasts(
  data[grepl("2", target)], 
  summarise_by = c("model", "target_type", "range", "quantile"), 
  compute_relative_skill = TRUE
)

@@

## Further interesting things

- look at crowd performance as a function of the number of forecasts --> maybe leave that to a second paper
